### 2024-07-25 Manuscript Analysis/manuscript_analysis.Rmd
WM pupil analysis

---
title: "Cognitive Analysis"
author: "Micah E. Hirsch"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

The purpose of this document is to share the analysis code and report the results for the manuscript. The full Rmd version of the document can be found on the project's Github repository or OSF project page. The data cleaning code for the pupil and perceived listening effort ratings can be found in the Github repo. Analyses were conducted using R version 4.3.3.

```{r}

# Loading in the needed packages

library(tidyverse)
library(rio)
library(glmmTMB)
library(janitor)
library(ggpubr)
library(patchwork)
library(gt)
library(gtsummary)
library(ggridges)
library(performance)
library(sjPlot)
library(ggcorrplot)
library(mgcv) # install.packages("mgcv")
library(itsadug) # install.packages("itsadug")

# Load in the listener demographics data

demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv") |>
  dplyr::filter(pupil_complete == "complete") |>
  dplyr::select(id, age, gender, ethnicity, race, native_lang) |>
  dplyr::mutate(gender = factor(gender, levels = c("Man", "Woman", "Nonbinary", "Questioning", "Prefer not to answer")),
                ethnicity = factor(ethnicity, levels = c("Hispanic/Latino(a/e)", "Not Hispanic/Latino(a/e)", "Prefer not to answer")),
                race = factor(race, levels = c("white/Caucasian", "Black/African American", "Asian/Asian American",
                                               "Native Hawaiian or Other Pacific Islander", "Native American or Alaska Native",
                                               "Biracial or Multiracial", "Race not listed", "Prefer not to answer")),
                native_lang = factor(native_lang, c("American English", "Not American English"))) |>
  dplyr::rename(subject = id)

# Loading in the cog data

cog <- rio::import("Cleaned Data/cleaned_cog_data.csv")

# Loading in the pupil data

pupil <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv")

# Loading in the phrase accuracy data

phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate")),
                rep_acc_rel = ifelse(rel_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_rel, "accurate", "inaccurate")))

```

# Listener Characteristics

## Demographic Information

The listeners used in this study are the same from the parent study. In brief, the demographic information from these listeners are below. A total of 34 young adult listeners participated in the experiment.

```{r}

demo_table <- demo |>
  dplyr::select(age, gender, ethnicity, race, native_lang) |>
  tbl_summary(type = list(age ~ "continuous",
                          gender ~ "categorical",
                          ethnicity ~ "categorical",
                          race ~ "categorical",
                          native_lang ~ "categorical"),
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(everything() ~ c(2)),
              label = list(age ~ "Age",
                           gender ~ "Gender",
                           ethnicity ~ "Ethnicity",
                           race ~ "Race",
                           native_lang ~ "Native Language")) |>
  as_gt()

demo_table

demo_table |>
  gt::gtsave("Tables/demo_table.html")

```

## Listener Cognition

The information below reports the mean age-corrected standardized scores on the subtests from the NIH Toolbox cognition battery (working memory, inhibitory control, cognitive flexibility, and processing speed).

```{r}

cog_table <- cog |>
  dplyr::select(-subject) |>
  # Just using the age-corrected scores
  dplyr::select(ends_with("_c")) |>
  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})"),
              digits = list(everything() ~ c(2)),
              label = list(list_sort_c ~ "Working Memory",
                           flanker_c ~ "Inhibitory Control",
                           card_sort_c ~ "Cognitive Flexibility",
                           pattern_c ~ "Processing Speed")) |>
  as_gt()

cog_table

cog_table |>
  gt::gtsave("Tables/cog_table.html")

```

```{r}

# Color Palette for Figures
my_pal <- c("#033F63", "#2CB5AE", "#28666E", "#4BAA9B", "#49E9B9", "#57A773")

cog |>
  dplyr::select(subject, ends_with("_c")) |>
  dplyr::rename("Working Memory" = list_sort_c,
                "Inhibitory Control" = flanker_c,
                "Cognitive Flexibility" = card_sort_c,
                "Processing Speed" = pattern_c) |>
  tidyr::pivot_longer(cols = 'Working Memory':'Processing Speed',
                      names_to = "cog_measure",
                      values_to = "score") |>
  ggplot() +
  aes(x = score,
      y = cog_measure,
      color = cog_measure,
      fill = cog_measure) +
  geom_density_ridges(jittered_points = T,
                      position = position_points_jitter(width = 0.05, height = 0),
                      point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.5) +
  labs(x = "Score", y = "", color = "Cognitive Measure", fill = "Cognitive Measure") +
  scale_color_manual(values = c(my_pal[3],my_pal[4], my_pal[5], my_pal[6])) +
  scale_fill_manual(values = c(my_pal[3],my_pal[4], my_pal[5], my_pal[6])) +
  theme_classic() +
  theme(legend.position = "none")

```

```{r}

# Removing unneeded items

rm(cog_table, demo_table, demo)

```

# Pupil Dilation

```{r}

# Selecting needed variables from phrase repetition accuracy df

acc <- phrase_acc |>
  dplyr::select(subject, trial, target_number, correct_words_initial, rep_acc)

## Merging acc df with pupil dilation df

pupil <- pupil |>
  # Merging phrase accuracy information with df
  dplyr::left_join(acc, by = c("subject", "trial")) |>
  # Filter out trials with missing phrase repetition data |>
  dplyr::filter(!is.na(rep_acc)) |>
  # Merging cognitive data
  dplyr::left_join(cog, by = "subject")

```

## Base Model

Final Model from parent study

```{r}

pupil <- pupil |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")),
                phrase = as.factor(code),
                listener = as.factor(subject),
                condition = case_when(speaker == "Control" & rep_acc == "accurate" ~ "Control Accurate",
                                      speaker == "Control" & rep_acc == "inaccurate" ~ "Control Inaccurate",
                                      speaker == "ALS" & rep_acc == "accurate" ~ "ALS Accurate",
                                      TRUE ~ "ALS Inaccurate"),
                condition = factor(condition, 
                                   levels = c("Control Accurate", "Control Inaccurate", "ALS Accurate", "ALS Inaccurate"))) |>
  dplyr::group_by(listener, speaker) |>
  dplyr::mutate(num_points = n()) |>
  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1)))) |>
  ungroup()

m0_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m = 1),
                discrete = T,
                AR.start = start_event,
                rho = .9341941,
                family = "scat",
                data = pupil)

summary.gam(m0_pupil)

m0_prediction <- plot_smooth(m0_pupil,
            view = "time_norm",
            plot_all = "condition",
            rug = F, se = 1, rm.ranef = T)

```

```{r}

m0_predicted_values <- m0_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
                speaker = factor(speaker, levels = c("Control", "ALS")))

m0_predicted_values |>
  ggplot() +
  aes(x = time_norm,
      y = fit,
      color = speaker,
      fill = speaker) +
  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
  geom_line() +
  geom_vline(xintercept = 0, size = 0.35) +
  geom_vline(xintercept = 1910, linetype = 2) +
  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
  facet_wrap("rep_acc") +
  theme_bw() +
  theme(aspect.ratio = 1,
        legend.position = "bottom") +
  labs(x = "Normalized Time (ms)",
       y = "Pupil Dilation (Arbitrary Units)",
       color = "Speaker",
       fill = "Speaker",
       linetype = "Speaker") +
  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS"))

```

## Working Memory

### Initial Model

```{r}

wm_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  te(time_norm, list_sort_c, k = 20) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m = 1),
                discrete = T,
                AR.start = start_event,
                rho = .9341941,
                family = "scat",
                data = pupil)

summary.gam(wm_pupil)

```

#### Model Check

```{r}

gam.check(wm_pupil)

acf_resid(wm_pupil)

```

#### Model Comparison

```{r}

compareML(m0_pupil, wm_pupil, signif.stars = T, suggest.report = T)

```

### Model 1

```{r}

wm_pupil_1 <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  te(time_norm, list_sort_c, by = speaker, k = 20) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m = 1),
                discrete = T,
                AR.start = start_event,
                rho = .9341941,
                family = "scat",
                data = pupil)

summary.gam(wm_pupil_1)

```

#### Model Check

```{r}

gam.check(wm_pupil_1)

acf_resid(wm_pupil_1)

```

#### Model Comparison

```{r}

compareML(m0_pupil, wm_pupil_1, signif.stars = T, suggest.report = T)

```
### Model 2

```{r}

wm_pupil_2 <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  te(time_norm, list_sort_c, by = condition, k = 20) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m = 1),
                discrete = T,
                AR.start = start_event,
                rho = .9341941,
                family = "scat",
                data = pupil)

summary.gam(wm_pupil_2)

```

#### Model Check

```{r}

gam.check(wm_pupil_2)

acf_resid(wm_pupil_2)

```
#### Model Comparison

```{r}

compareML(m0_pupil, wm_pupil_2, signif.stars = T, suggest.report = T)

```