### 2024-10-21 Manuscript Analysis/ASHA Poster Analyses.Rmd
first run through of asha analyses

---
title: "Poster Analysis"
author: "Micah Hirsch"
date: "2024-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document provides the code for the peak pupil dilation analysis for a poster presentation presented at the 2024 Annual Speech Language and Hearing Association Convention.

```{r}

# Loading in the needed packages

library(tidyverse) # install.packages("tidyverse")
library(rio) # install.packages("rio")
library(glmmTMB) # install.packages("glmmTMB")
library(janitor) # install.packages("janitor")
library(ggpubr) # install.packages("ggpubr")
library(patchwork) # install.packages("patchwork")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(ggridges) # install.packages("ggridges")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")
library(ggcorrplot) # install.packages("ggcorrplot")
library(see) #install.packages("see")

# Loading in the cog data

cog <- rio::import("Cleaned Data/cleaned_cog_data.csv")

# Loading in the pupil data

pupil <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv")

# Loading in the phrase accuracy data

phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate")),
                rep_acc_rel = ifelse(rel_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_rel, "accurate", "inaccurate"))) |>
  dplyr::select(subject, trial, rep_acc)

```

# Peak Pupil Dilation

```{r}

peak_pupil_df <-pupil |>
  dplyr::group_by(subject, trial, speaker, code) |>
  dplyr::summarize(peak_pupil = max(normed_pupil)) |>
  dplyr::ungroup() |>
  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
  dplyr::left_join(cog, by = "subject") |>
  dplyr::mutate(trial_c = trial - 6,
                speaker = factor(speaker, levels = c("Control", "ALS")),
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate"))) |>
  dplyr::filter(!is.na(rep_acc))

```
## Visualization

```{r}

# Color Palette for Figures
my_pal <- c("#25CED1", "#0D3B66", "#F75C03", "#3B5249", "#F1C40F", "#D90368")

peak_pupil_df |>
  dplyr::select(subject, trial, speaker, peak_pupil, rep_acc, ends_with("_c")) |>
  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
  dplyr::summarize(peak_pupil = mean(peak_pupil)) |>
  dplyr::rename('Working Memory' = list_sort_c,
                'Inhibitory Control' = flanker_c,
                'Cognitive Flexibility' = card_sort_c,
                'Processing Speed' = pattern_c) |>
  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
                     names_to = "cog_measure",
                     values_to = "score") |>
  ungroup() |>
  ggplot() +
  aes(x = score,
      y = peak_pupil,
      color = speaker,
      shape = speaker) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm") +
  facet_grid(rep_acc ~ cog_measure) +
  labs(x = "Score", y = "Peak Pupil Dilation") +
  theme_classic() +
  scale_color_manual(values = my_pal) +
  scale_shape_manual(values = c(1, 2)) +
  theme(aspect.ratio = 1)

```
## Models

### Fully Unconditional Model

```{r}

m0 <- glmmTMB(peak_pupil ~ 1 + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m0, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")

performance::icc(m0)
performance::check_model(m0)

```

```{r}

m1 <- glmmTMB(peak_pupil ~ 1 + (trial_c|subject), data = peak_pupil_df)
sjPlot::tab_model(m1, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m0, m1)

```

```{r}

m2 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = peak_pupil_df)
sjPlot::tab_model(m2, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m1, m2)

```

```{r}

m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m3, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m2, m3)

```
```{r}

m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m3b, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3, m3b)

```

```{r}

m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m4, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m4)

```

```{r}

m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m5, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m5)

```

```{r}

m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(pattern_c,scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m6, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m6)

```

```{r}

m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m7, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m7)

```

```{r}

m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m8, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m7, m8)

```
```{r}

m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m9, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m7, m9)

```

```{r}

m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + speaker*scale(pattern_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m10, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m9, m10)

```

```{r}

m11 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m11, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m9, m11)

```

```{r}

m12 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(flanker_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m12, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m11, m12)

```

```{r}

m13 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m13, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m11, m13)

```

```{r}

m14 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m14, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m11, m14)

```

```{r}

m15 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(list_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m15, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m15)

```

```{r}

m16 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(flanker_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m16, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m16)

```

```{r}

m17 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(card_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m17, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m17)

```

```{r}

m18 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(pattern_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m18, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m18)

```

```{r}

m14b <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                  rep_acc*scale(pattern_c, scale = F) +
                  (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m14b, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::check_model(m14b)

```